---
title: "Project 3"
author: "John Mazon"
date: "9/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### MYSQL Database integration with R


#### Additional packages installed
Please note below additional packages installed. The formattable package is used to transform vectors and data frames into more readable and impactful tabular formats. 
Package 'conflicted' solved an issue with the rename func as it pertains to knit/html.

install.packages("formattable")
install.packages("conflicted")

```{r}
library(tidyverse)
library(RMySQL)
library(formattable)

```


```{r}
##DBconnect was utilized in this assignment to authenticate/pull in the flight delay information into R
my.database10 = dbConnect(MySQL(), user='root', password = 'Password1', dbname='jobs', host='localhost')
dbListTables(my.database10)

generalskills_list <- fetch(dbSendQuery(my.database10, "SELECT * FROM general_skills"))
dim(generalskills_list)
colnames(generalskills_list)
summary(generalskills_list)

```



```{r}
##DBconnect was utilized in this assignment to authenticate/pull in the flight delay information into R
my.database10 = dbConnect(MySQL(), user='root', password = 'Password1', dbname='jobs', host='localhost')
dbListTables(my.database10)

indeed_list <- fetch(dbSendQuery(my.database10, "SELECT * FROM indeed_jobs"))
dim(indeed_list)
colnames(indeed_list)

```

Experimented with the tibble func below for the first time. Tibbles have a refined print method that shows only the first 10 rows, and all the columns that fit on screen. This makes it much easier to work with large data.
```{r}
library(tidyverse)
library(dplyr)


indeed_listinfo <- as_tibble(indeed_list)
indeed_listinfo

generalskills_listinfo <- as_tibble(generalskills_list)
generalskills_listinfo

indeed_listinfo2 <- data.frame(indeed_listinfo$MyUnknownColumn,
                               indeed_listinfo$Job_Title,
                               indeed_listinfo$Job_Type,
                               indeed_listinfo$Skill,
                               indeed_listinfo$No_of_Skills,
                               indeed_listinfo$Description,
                               indeed_listinfo$Location,
                               indeed_listinfo$Company_Industry,
                               indeed_listinfo$python,
                               indeed_listinfo$sql,
                               indeed_listinfo$r,
                               indeed_listinfo$hadoop,
                               indeed_listinfo$tableau,
                               indeed_listinfo$sas,
                               indeed_listinfo$spark,
                               indeed_listinfo$java,
                               indeed_listinfo$Others,
                               indeed_listinfo$CA,
                               indeed_listinfo$NY,
                               indeed_listinfo$VA,
                               indeed_listinfo$TX,
                               indeed_listinfo$MA,
                               indeed_listinfo$IL,
                               indeed_listinfo$WA,
                               indeed_listinfo$MD,
                               indeed_listinfo$DC,
                               indeed_listinfo$NC,
                               indeed_listinfo$Other_states,
                               indeed_listinfo$`Consulting and Business Services`,
                               indeed_listinfo$`Internet and Software`,
                               indeed_listinfo$`Banks and Financial Services`,
                               indeed_listinfo$`Health Care`,
                               indeed_listinfo$Insurance,
                               indeed_listinfo$Other_industries)

(colnames(indeed_listinfo2))
```


```{r}
library(dplyr)

indeed_listinfo3 <- dplyr::rename(indeed_listinfo2,
          "ID"="indeed_listinfo.MyUnknownColumn",
          "Title" = "indeed_listinfo.Job_Title",
          "Type" = "indeed_listinfo.Job_Type",
          "Skill" = "indeed_listinfo.Skill",
          "Skills" = "indeed_listinfo.No_of_Skills",
          "Description" = "indeed_listinfo.Description",
          "Location" = "indeed_listinfo.Location",
          "Industry" = "indeed_listinfo.Company_Industry",
          "Python" = "indeed_listinfo.python",
          "SQL" = "indeed_listinfo.sql",
          "R" = "indeed_listinfo.r",
          "Hadoop" = "indeed_listinfo.hadoop",
          "Tableau" = "indeed_listinfo.tableau",
          "SAS" = "indeed_listinfo.sas",
          "Spark" = "indeed_listinfo.spark",
          "Java" = "indeed_listinfo.java",
          "Others" = "indeed_listinfo.Others",
          "CA" = "indeed_listinfo.CA",
          "NY" = "indeed_listinfo.NY",
          "VA" = "indeed_listinfo.VA",
          "TX" = "indeed_listinfo.TX",
          "MA" = "indeed_listinfo.MA",
          "IL" = "indeed_listinfo.IL",
          "WA" = "indeed_listinfo.WA",
          "MD" = "indeed_listinfo.MD",
          "DC" = "indeed_listinfo.DC",
          "NC" = "indeed_listinfo.NC",
          "Other" = "indeed_listinfo.Other_states",
          "Consulting" = "indeed_listinfo..Consulting.and.Business.Services.",
          "Internet" = "indeed_listinfo..Internet.and.Software.",
          "Financial" = "indeed_listinfo..Banks.and.Financial.Services.",
          "Healthcare" = "indeed_listinfo..Health.Care.",
          "Insurance" = "indeed_listinfo.Insurance",
          "Varied" = "indeed_listinfo.Other_industries"

)



table(indeed_listinfo3$Python)
table(indeed_listinfo3$SQL)
table(indeed_listinfo3$R)
table(indeed_listinfo3$Hadoop)
table(indeed_listinfo3$Tableau)
table(indeed_listinfo3$SAS)
table(indeed_listinfo3$Spark)
table(indeed_listinfo3$Java)
table(indeed_listinfo3$Others)

library(data.table)

setDT(indeed_listinfo3)[, .N, by=Python]
setDT(indeed_listinfo3)[, .N, by=SQL]
setDT(indeed_listinfo3)[, .N, by=R]
setDT(indeed_listinfo3)[, .N, by=Hadoop]
setDT(indeed_listinfo3)[, .N, by=Tableau]
setDT(indeed_listinfo3)[, .N, by=SAS]
setDT(indeed_listinfo3)[, .N, by=Spark]
setDT(indeed_listinfo3)[, .N, by=Python]
setDT(indeed_listinfo3)[, .N, by=Java]




```



















